Assessment of zonally symmetric and asymmetric components of the Southern Annular Mode using a novel approach

### 2021-05-06 analysis/paper/asymsam.Rmd
Achica espacio entre plots polares
Closes #29

```{r method, fig.cap = "Spatial patterns of the first EOF of 700~hPa geopotential height for 1979 -– 2018 period. (a) Full field, (b) zonally asymmetric component and (c) zonally symmetric component. Arbitrary units; positive values in blue and negative values in red.", fig.width=6, fig.height=2.5, fig.env="figure*"}
sam_sep <- SAM[, eof[[1]]$right, by = .(lev)] %>% 
  setnames("hgt", "full") %>% 
  .[, c("sym", "asym") := list(mean(full), Anomaly(full)), by = .(lat, lev)]

@@

sam_plot %>% 
  periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +
  # geom_contour(aes(z = value, linetype = factor(-sign(..level..))), global.breaks = FALSE) +
  geom_contour_fill(aes(z = value), global.breaks = FALSE, breaks = ZeroBreaks) +
  geom_contour_tanaka2(aes(z = value), global.breaks = FALSE, breaks = ZeroBreaks) +
  geom_qmap() +
  geom_coords() +
  scale_x_longitude(labels = NULL) +
  scale_y_latitude(limits = c(NA, -20), labels = NULL) +
  scale_fill_divergent(guide = "none", high = "#731C1F", low = "#323071") +
  coord_polar() +
  facet_grid(.~variable, labeller = labeller(variable = label_parsed, lev = lev.lab)) +
  axis_labs_smol +
  no_grid +
  tag_facets()
```

@@

```{r 2d-regr, fig.cap = "Regression of geopotential height (meters) at (row 1) 50~hPa and (row 2) 700~hPa with (column a) SAM, (column b) A\\nobreakdash-SAM, and (column c) S\\nobreakdash-SAM for the 1979 -- 2018 period. The regression patterns for A\\nobreakdash-SAM and S\\nobreakdash-SAM are the result of one multiple regression using both indices. Points marked on panel b.2 are the location of the reference points used by \\cite{raphael2004} for their Zonal Wave 3 index. ", fig.env = "figure*", fig.width=6, fig.height=5}
indices_2d_yearly %>% 
  .[lat <= -20] %>% 
  .[lev %in% c(50, 700)] %>%
  .[, term := factor(term, levels = c("full", "asym", "sym"))] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = AnchorBreaks(0, exclude =  0),
                    global.breaks = FALSE) +
  geom_contour_tanaka2(aes(z = estimate),
                      breaks = AnchorBreaks(0, exclude =  0),
                      size = 0.3,
                      global.breaks = FALSE) +
  # geom_contour2(aes(z = env)) +
  geom_qmap(~.x[lat <= -20]) +
  geom_coords() +
  geom_point(data = data.frame(lon = ConvertLongitude(c(50, 166, -76)),
                               lat = -49, term = factor("asym", 
                                                        levels = c("full", "asym", "sym")),
                               lev = 700),
             shape = 21, size = 2.5, color = "black", fill = "#de3e80") +
  scale_x_longitude(labels = NULL) +
  scale_y_latitude(limits = c(NA, -20), labels = NULL) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-45, 45),
                       oob = scales::squish,
                       breaks = AnchorBreaks(0, 10, 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(lev ~ term, labeller = labeller(lev = lev.lab, term = lab_sam)) +
  tag_facets("cr") +
  no_grid 
```

@@

  geom_qmap(~.x[lat <= 0]) +
  geom_coords() +
  scale_x_longitude(labels = NULL) +
  scale_y_latitude(labels = NULL, limits = c(NA, 0)) +
  discrete_scale("fill", "name", name = "2-meter air temperature",
                 palette = discrete_div,
                 guide = guide_colorsteps_bottom(show.limits = TRUE)) +
  coord_polar() +
  facet_grid(season ~ term, labeller = labeller(term = lab_sam, lev = lev.lab)) +
  tag_facets("cr", tag_suffix = "") +
  no_grid 
```

@@

lons <- c(0, unique(pp_regr$lon), 360)
pp_regr %>%
  .[season == "Annual"] %>%
  .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(season, term)] %>% 
  # Need to interpolate 0 and 360 to get closed contours
  qwrap(lon = c(1.25, 360 + 1.25) ~ c(-1.25, 360 + 1.25)) %>% 
  .[, lapply(.SD, function(x) approx(lon, x, xout = lons)$y), by = .(lat, season, term)] %>%
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate, fill = ..level..), 
                    breaks = breaks_pp) +
  stat_subset(aes(subset = is.na(estimate)), geom = "tile", color = "gray50") +
  geom_contour2(aes(z = p.value), breaks = 0.05, size = 0.2) +
  geom_qmap(~.x[lat <= 10], keep = 0.015, weighting = 0.9) +
  scale_x_longitude(breaks = NULL) +
  scale_y_latitude(breaks = NULL, limits = c(-90, NA)) +
  discrete_scale("fill", "name", name = "Precipitation",
                 palette = discrete_div_pp,
                 guide = guide_colorsteps_bottom(show.limits = TRUE)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  geom_coords() +
  # coord_quickmap(ylim = c(NA, 0)) +
  facet_grid(.~term, labeller = labeller(term = lab_sam)) +
  tag_facets("panel", position = "tl")  +
  no_grid 
```

>>>>

```{r method, fig.cap = "Spatial patterns of the first EOF of 700~hPa geopotential height for 1979 -– 2018 period. (a) Full field, (b) zonally asymmetric component and (c) zonally symmetric component. Arbitrary units; positive values in blue and negative values in red.", fig.width=6, fig.height=2.7, fig.env="figure*"}
sam_sep <- SAM[, eof[[1]]$right, by = .(lev)] %>% 
  setnames("hgt", "full") %>% 
  .[, c("sym", "asym") := list(mean(full), Anomaly(full)), by = .(lat, lev)]

@@

sam_plot %>% 
  periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +
  # geom_contour(aes(z = value, linetype = factor(-sign(..level..))), global.breaks = FALSE) +
  geom_contour_fill(aes(z = value), global.breaks = FALSE, breaks = ZeroBreaks) +
  geom_contour_tanaka2(aes(z = value), global.breaks = FALSE, breaks = ZeroBreaks) +
  geom_qmap() +
  geom_coords() +
  scale_x_longitude(labels = NULL) +
  scale_y_latitude(limits = c(NA, -20), labels = NULL) +
  scale_fill_divergent(guide = "none", high = "#731C1F", low = "#323071") +
  coord_polar() +
  facet_grid(.~variable, labeller = labeller(variable = label_parsed, lev = lev.lab)) +
  axis_labs_smol +
  no_grid +
  tag_facets() +
  theme(panel.spacing = grid::unit(-1, "lines"))
```

@@

```{r 2d-regr, fig.cap = "Regression of geopotential height (meters) at (row 1) 50~hPa and (row 2) 700~hPa with (column a) SAM, (column b) A\\nobreakdash-SAM, and (column c) S\\nobreakdash-SAM for the 1979 -- 2018 period. The regression patterns for A\\nobreakdash-SAM and S\\nobreakdash-SAM are the result of one multiple regression using both indices. Points marked on panel b.2 are the location of the reference points used by \\cite{raphael2004} for their Zonal Wave 3 index. ", fig.env = "figure*", fig.width=6, fig.height=5}
indices_2d_yearly %>% 
  .[lat <= -20] %>% 
  .[lev %in% c(50, 700)] %>%
  .[, term := factor(term, levels = c("full", "asym", "sym"))] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = AnchorBreaks(0, exclude =  0),
                    global.breaks = FALSE) +
  geom_contour_tanaka2(aes(z = estimate),
                      breaks = AnchorBreaks(0, exclude =  0),
                      size = 0.3,
                      global.breaks = FALSE) +
  # geom_contour2(aes(z = env)) +
  geom_qmap(~.x[lat <= -20]) +
  geom_coords() +
  geom_point(data = data.frame(lon = ConvertLongitude(c(50, 166, -76)),
                               lat = -49, term = factor("asym", 
                                                        levels = c("full", "asym", "sym")),
                               lev = 700),
             shape = 21, size = 2.5, color = "black", fill = "#de3e80") +
  scale_x_longitude(labels = NULL) +
  scale_y_latitude(limits = c(NA, -20), labels = NULL) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-45, 45),
                       oob = scales::squish,
                       breaks = AnchorBreaks(0, 10, 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(lev ~ term, labeller = labeller(lev = lev.lab, term = lab_sam)) +
  tag_facets("cr", position = list(x = 0.15, y = 0.85)) +
  no_grid +
  theme(panel.spacing = grid::unit(-1, "lines"))
```

@@

  geom_qmap(~.x[lat <= 0]) +
  geom_coords() +
  scale_x_longitude(labels = NULL) +
  scale_y_latitude(labels = NULL, limits = c(NA, 0)) +
  discrete_scale("fill", "name", name = "2-meter air temperature",
                 palette = discrete_div,
                 guide = guide_colorsteps_bottom(show.limits = TRUE)) +
  coord_polar() +
  facet_grid(season ~ term, labeller = labeller(term = lab_sam, lev = lev.lab)) +
  tag_facets("cr", position = list(x = 0.15, y = 0.85)) +
  no_grid +
  theme(panel.spacing = grid::unit(-1, "lines"))
```

@@

lons <- c(0, unique(pp_regr$lon), 360)
pp_regr %>%
  .[season == "Annual"] %>%
  .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(season, term)] %>% 
  # Need to interpolate 0 and 360 to get closed contours
  qwrap(lon = c(1.25, 360 + 1.25) ~ c(-1.25, 360 + 1.25)) %>% 
  .[, lapply(.SD, function(x) approx(lon, x, xout = lons)$y), by = .(lat, season, term)] %>%
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate, fill = ..level..), 
                    breaks = breaks_pp) +
  stat_subset(aes(subset = is.na(estimate)), geom = "tile", color = "gray50") +
  geom_contour2(aes(z = p.value), breaks = 0.05, size = 0.2) +
  geom_qmap(~.x[lat <= 0], keep = 0.015, weighting = 0.9) +
  scale_x_longitude(breaks = NULL) +
  scale_y_latitude(breaks = NULL, limits = c(-90, NA)) +
  discrete_scale("fill", "name", name = "Precipitation",
                 palette = discrete_div_pp,
                 guide = guide_colorsteps_bottom(show.limits = TRUE)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  geom_coords() +
  # coord_quickmap(ylim = c(NA, 0)) +
  facet_grid(.~term, labeller = labeller(term = lab_sam)) +
  # tag_facets("panel", position = "tl")  +
  tag_facets("panel", position = list(x = 0.15, y = 0.85)) +
  no_grid +
  theme(panel.spacing = grid::unit(-1, "lines"))
```
